Comparison of Catalogs

### 2024-10-05 compare/compare.qmd
def and expl

In this script we will compare 2 catalogs @kovlakasHeraklionExtragalacticCatalogue2021 and [@karachentsevUPDATEDNEARBYGALAXY2013, @karachentsevSTARFORMATIONPROPERTIES2013a]

@@

```{python}
#| echo: false
import numpy as np
import matplotlib.pyplot as plt
from sklearn.metrics import r2_score
import astropy
from astropy.io import ascii
from astropy.coordinates import SkyCoord
from astropy.table import Table, join, QTable
import astropy.units as u
from astropy.visualization import quantity_support, hist
import scipy
#quantity_support()
from astropy.stats import sigma_clip, SigmaClip
from astropy.modeling import models, fitting
from scipy.stats import pearsonr
import seaborn as sns
import pandas as pd
import glob
import os
from tabulate import tabulate
from IPython.display import Markdown, display

@@

Checking for completeness in galaxy catalogs is essential to ensure that the data accurately represents the true population of galaxies. Incomplete catalogs can lead to biased results in statistical studies, such as the distribution of galaxy luminosity, mass, or star formation rates. Additionally, missing galaxies, especially those at faint magnitudes or large distances, can distort cosmological measurements and hinder our understanding of galaxy formation and evolution.

@@

2. Distances of D\<11 Mpc

@@

-   Completeness within a 10 Mpc radius is difficult to assess due to:
    -   Variability in galaxy properties (luminosity, size, surface brightness, gas content)
    -   Errors in distance measurements (Tully–Fisher method errors of \~20–25%), especially at the 10 Mpc boundary.
        -   Accurate distances are mainly known within \~5 Mpc.
    -   Non-Hubble motions (\~300 km/s) may make up half of the adopted velocity constraint @eq-lcv-vel
        -   Solution for our usage: only keep the galaxies inside a radius of $D=11 Mpc$
        -   HI sources in surveys with low angular resolution: "The presence around our Galaxy of hundreds of high-velocity clouds with low line-of-sight velocities and small W50 widths also provokes the inclusion of false “nearby” dwarf galaxies in the LV"@karachentsevUPDATEDNEARBYGALAXY2013
        -   photographic emulsion defect -\> exotic cases of galaxies -\> radial velocity of $+614 km s^{−1}$
-   Astro-Spam and Survey Errors:
    -   Automatic surveys produce false detections (e.g., stars misclassified as galaxies, high-velocity clouds mistaken for dwarfs).
-   Conditional Completeness Estimate:
    -   Galaxies brighter than $M_B^c = -11^m$ or with linear diameters larger than $A_{26} = 1.0$ kpc show (40–60)% completeness.
    -   "among the members of the Local Group (D \< 1 Mpc), only half of the galaxies have absolute magnitudes brighter than −11m. Consequently, more than half of the ultra-faint dwarf companions around normal galaxies, like the Sombrero galaxy (D = 9.3 Mpc), still remain outside our field of view."
-   Undetected Ultra-Faint Dwarfs:
    -   Many faint dwarf galaxies remain undetected beyond the Local Group. Estimated population of undetected dwarfs could be as large as $10^3–10^4$ within the LV.
-   Surface Brightness Distribution:
    -   Surface brightness remains consistent across distances, except for ultra-faint dwarfs ($SB <31 \text{mag}\cdot\text{arcsec}^{-2}$).
    -   Faintest dwarf galaxies are detectable only nearby due to their resolution into individual stars.
-   Luminosity-Size Relationship:
    -   Observations align with cosmological models predicting $L\sim A^3$ @navarroStructureColdDark1996, though deviations occur for extremely low surface brightness galaxies.
    -   "The deviation from it at the extremely low surface brightness end is due to a systematic overestimation of dwarf galaxy sizes, the brightness profiles of which lie entirely below the Holmberg isophote."

@@

```{python}
def plot_completeness(dataset_choice="lcv", column_name="Dis", description="Distance", units=True, logx=False, logy=False, bins=0, keep_nan=False):
    """
    Plots histograms with KDE and displays a table of counts and percentages for each bin, 
    based on the choice of dataset (LCV or HEC).

    Parameters:
    - dataset_choice: A string indicating which dataset to use ("lcv" or "hec")
    - column_name: The column name in the Qtable to plot and analyze (default is "Dis")
    - description: What to show in the title and the axis (default "Distance")
    - units: Show units or not (default is True)
    - logx: logarithmic x-axis (default is False)
    - logy: logarithmic y-axis (default is False)
    - bins: Number of bins or custom bin edges (default is 'auto')
    - keep_nan: keep (True) or remove NaN's (default is True)
    """

    if dataset_choice == "lcv":
        main_df, unique_df, label = lcv, lcv_not_hec, "LCV"
    elif dataset_choice == "hec" and hec is not None and hec_not_lcv is not None:
        main_df, unique_df, label = hec, hec_not_lcv, "HECATE"
    else:
        raise ValueError("Invalid dataset choice. Please choose 'lcv' or 'hec' with valid datasets.")

    # Remove NaN values before plotting
    if not keep_nan:
        main_data_clean = main_df[~np.isnan(main_df[column_name].data)].data
        unique_ddata_clean = unique_df[~np.isnan(unique_df[column_name].data)]
    else:
        main_data_clean = main_df[column_name].data
        unique_data_clean = unique_df[column_name].data

    # Check if bins is an integer or array of bin edges
    if bins == 0:
        counts, bin_edges = np.histogram(main_data_clean, bins='auto')
    elif isinstance(bins, int):
        counts, bin_edges = np.histogram(main_data_clean, bins=bins)
    else:
        bin_edges = bins  # If bins is an array of bin edges

    # Plot the histograms
    sns.histplot(main_data_clean, bins=bin_edges, kde=True, label=label)
    sns.histplot(unique_data_clean, bins=bin_edges, kde=True, label=f"Unique for {label}")

    # Set logarithmic axes if requested
    if logx:
        plt.xscale('log')
    if logy:
        plt.yscale('log')

    # Set the title and labels
    plt.title(f"Completeness of {description} for {label}")
    if units and main_df[column_name].unit is not None:
        plt.xlabel(f"{description} [{main_df[column_name].unit}]")
    else:
        plt.xlabel(f"{description}")
    plt.ylabel("Number of galaxies")
    plt.legend()

    # Calculate histogram counts for each dataset
    main_counts, _ = np.histogram(main_data_clean, bins=bin_edges)
    unique_counts, _ = np.histogram(unique_data_clean, bins=bin_edges)

    # Calculate bin centers for annotation
    bin_centers = 0.5 * (bin_edges[1:] + bin_edges[:-1])

    # Annotate percentages on the bars
    for i, center in enumerate(bin_centers):
        if main_counts[i] > 0:
            unique_pct = (unique_counts[i] / main_counts[i]) * 100
            plt.text(center, unique_counts[i], f'{unique_pct:.1f}%', color="green", ha='center', va='bottom')

    # Show the plot
    plt.show()
    plt.close()

@@

### Distance

@@

plt.subplot(2,1,1)
sns.histplot(main, bins = bin_edges, kde = True, label = "HECATE")
sns.histplot(unique, bins = bin_edges, kde = True, label = "Unique for HECATE")
plt.legend()
plt.title("Completenes of Distance for HECATE")
unique_pct = (unique_counts/main_counts)*100 
###
plt.subplot(2,1,2)
ax = sns.lineplot(x = bin_edges[1:], y = unique_pct)
ax.set(xlabel='xlabel', ylabel='ylabel', title='title')

#changing ylables ticks
plt.yticks(np.arange(min(unique_pct), max(unique_pct)+1, 5.0))
y_value=['{:,.0f}'.format(x) + '%' for x in ax.get_yticks()]
ax.set_yticklabels(y_value)
plt.ylabel("Completeness %")
plt.xlabel("Distance [{}]".format(hec["D"].unit))
plt.title("")
plt.show()

@@

plt.subplot(2,1,1)
sns.histplot(main, bins = bin_edges, kde = True, label = "LCV")
sns.histplot(unique, bins = bin_edges, kde = True, label = "Unique for LCV")
plt.legend()
plt.title("Completenes of Distance for LCV")
unique_pct = (unique_counts/main_counts)*100 
###
plt.subplot(2,1,2)
ax = sns.lineplot(x = bin_edges[1:], y = unique_pct)
ax.set(xlabel='xlabel', ylabel='ylabel', title='title')

#changing ylables ticks
plt.yticks(np.arange(min(unique_pct), max(unique_pct)+1, 5.0))
y_value=['{:,.0f}'.format(x) + '%' for x in ax.get_yticks()]
ax.set_yticklabels(y_value)

plt.ylabel("Completeness %")
plt.xlabel("Distance [{}]".format(lcv["Dis"].unit))
plt.title("")
plt.show()

@@

plt.subplot(2,1,1)
sns.histplot(main, bins = bin_edges, kde = True, label = "HECATE")
sns.histplot(unique, bins = bin_edges, kde = True, label = "Unique for HECATE")
plt.legend()
plt.title("Completenes of Types for HECATE")
unique_pct = (unique_counts/main_counts)*100 
###
plt.subplot(2,1,2)
ax = sns.lineplot(x = bin_edges[1:], y = unique_pct)
ax.set(xlabel='xlabel', ylabel='ylabel', title='title')

#changing ylables ticks
plt.yticks(np.arange(min(unique_pct), max(unique_pct)+1, 5.0))
y_value=['{:,.0f}'.format(x) + '%' for x in ax.get_yticks()]
ax.set_yticklabels(y_value)
plt.ylabel("Completeness %")
plt.xlabel("Types of galaxies")
plt.title("")
plt.show()

#LCV
main = lcv["TType"].data
main = main[~np.isnan(main)]
unique = lcv_not_hec["TType"].data
unique = unique[~np.isnan(unique)]
counts, bin_edges = np.histogram(main)

@@

plt.subplot(2,1,1)
sns.histplot(main, bins = bin_edges, kde = True, label = "LCV")
sns.histplot(unique, bins = bin_edges, kde = True, label = "Unique for LCV")
plt.legend()
plt.title("Completenes of Distance for LCV")
unique_pct = (unique_counts/main_counts)*100

plt.subplot(2,1,2)
ax = sns.lineplot(x = bin_edges[1:] ,y = unique_pct)
ax.set(xlabel='xlabel', ylabel='ylabel', title='title')

#changing ylables ticks
plt.yticks(np.arange(min(unique_pct), max(unique_pct)+1, 10.0))
y_value=['{:,.0f}'.format(x) + '%' for x in ax.get_yticks()]
ax.set_yticklabels(y_value)

plt.ylabel("Completeness %")
plt.xlabel("Types of galaxies")
plt.title("")
plt.show()
```

As we can see from the histograms @fig-dis-comp and @fig-type-comp the sample of unique galaxies of each catalog, gets smaller by an almost constant proportion (Inner join).

@@

## Scatter plots and $R^2$ calculation

1.  $R^2$: Measures the proportion of variance explained by the linear model.
2.  Slope of the Fitted Line: Should be close to 1 for a 1-1 correlation.[^1]

[^1]: Some data seem to have a very good linear correlation but they have many outliers. This is why we will clip the outliers with $\sigma > 3$

<!-- -->

@@

4.  [Plots]{.underline}: Plots are essential for visually assessing the relationship between two datasets, identifying correlations, trends, and outliers, and evaluating the fit of linear models.

-   Histograms: Because not all of our data have the same number of counts, the comparison with histograms between data that are not the same, doesn't help us right now.[^3] This is why we will only use histograms for comparing the distribution of same-data columns normalized by their maximum value

[^3]: When we will use the outer join table we could use histograms due to the large number of counts.

<!-- -->

-   Correlation Heatmaps: A correlation heatmap is a graphical tool that displays the correlation between multiple variables as a color-coded matrix. It's like a color chart that shows us how closely related different variables are. In a correlation heatmap, each variable is represented by a row and a column, and the cells show the correlation between them. The color of each cell represents the strength and direction of the correlation, with darker colors indicating stronger correlations.

@@

-   Kernel Density Estimate (KDE) plot: The KDE plot visually represents the distribution of data, providing insights into its shape, central tendency, and spread.

5.  Percentage change: We can calculate the percentage change of the data for each galaxy and then we can see if the data are similar, based on minimum, the maximum and the mean value of the difference.

@@

|  LCV  | HECATE |   Description   | Pearson Correlation \[-1,1\] |
|:-----:|:------:|:---------------:|:----------------------------:|
|  Dis  |   D    |    Distance     |      `{python} round(compare_data(dt["Dis"],dt["D"])[3],3)`       |

@@

- The average error of the distance in the HECATE catalog is $\overline{E_D} = \pm 1.6$ Mpc, so the intercept is included in the error.
- So we can assume that the Distances are the same

@@

- The average error of the radial velocity in the HECATE catalog is $\overline{E_V} = \pm 12\text{km}\cdot s^{-1}$ , so the intercept is included in the error.
- So we can assume that the radial velocities are the same

@@

```{python}
pairplot_with_correlation(vel_data)
```

@@

@@

@@

types = dt[["T", "TType","E_T"]]
types["diff"] = dt["T"]-dt["TType"]
types = types[np.abs(types["diff"])<7]

@@

"Morphological type of galaxy in the numerical code according to the classification by de Vaucouleurs et al. (1991). It should be noted that about three quarters of objects in the LV are dwarf galaxies, which require a more detailed morphological classification. For example, dwarf spheroidal galaxies and normal ellipticals are usually denoted by the same numerical code T \< 0, although their physical properties drastically differ. The classification problem arises as well for the “transient” type dwarf galaxies, T r, which combine the features of spheroidal (Sph) and irregular (Ir)systems. Due to small classification errors, such objects may “jump” from one end of the T scale to the other." @karachentsevUPDATEDNEARBYGALAXY2013

-   we can assume that the galaxies from the upper left and lower right regions of the plot are classified differently because of this
-   this might explain the large errors E_T of HECATE
-   why we see a difference in the distribution of @fig-type-comp

We can remove the "problematic" galaxies by only keeping the ones with: 
$$
|{T_{HECATE}-T_{LCV}}|<7
$$

- The average uncertainty of the morphological type in the HECATE catalog is $\overline{E_T} = \pm 1.4$, so the intercept is included in the error.
- So we can assume that the Distances are the same

<!-- #### Normalize the scale of galaxy types

It is very possible that the two catalogs use different scaling methods, as indicated by the use of decimal numbers in HECATE.

```{python}
types = dt[["T", "TType"]].to_pandas()
```

```{python}
def min_max_normalize(series):
    return (series - series.min()) / (series.max() - series.min())

types["T_norm"] = min_max_normalize(types["T"])
types["T_norm"].mask = types["T"].mask
types["TType_norm"] = min_max_normalize(types["TType"])
#types.describe()
```
TType

Also, as we can see the minimum values are lower by 2 in HECATE, which complies with the linear fit.

```{python}
sns.histplot(types[["TType_norm", "T_norm"]], kde = True)
plt.xlabel("Types of galaxies")
plt.show()
```

```{python}
types["norm_diff"] = (-types["TType_norm"] + types["T_norm"])
sns.histplot(types["norm_diff"], kde = True)

plt.xlabel(r"$T_{HECATE}-T_{LCV}$")
plt.show()
```

```{python}
types["norm_rel_diff"] = (-types["TType_norm"] + types["T_norm"])/types["T"]
sns.histplot(types["norm_rel_diff"], kde = True)
plt.xlim(-0.2,0.2)
plt.xlabel(r"$(T_{HECATE}-T_{LCV})/T_{HECATE}$")
plt.show()
```
-->

@@

```{python}
#| layout: [[50,50]]
plt.close()
temp = dt[["inc", "INCL"]].to_pandas()

@@

temp["Percentage Change [%]"] = (-temp["inc"] + temp["INCL"])/temp["INCL"]
temp.loc[np.isinf(temp["Percentage Change [%]"]), "Percentage Change [%]"] = np.nan
sns.histplot(temp["Percentage Change [%]"], kde = True)
plt.show()
```
```{python}
temp.describe()
```

We can see that for values in the range $[\sim 30^\circ,\sim 80^\circ]$, the values of the LCV inclination are higher. However, since their means, median, min and maxes are similar and the percentage change is practically 0% (mean, median, $\sigma$ = 0 with a range $[-3\%,1\%]$), we can ignore the differences and assume they are the same values.

@@

it is not very clear if we truly have a correlation or not. We need to see the linear correlation of the decimal logarithms.

$\overline{R_1} = 3.9$ [arcmin],  $\overline{a_{26}} = 9.3$ [arcmin], so the intercept is negligable.
$$
R_1 = 0.48\cdot a_{26}-0.34 \sim \frac{1}{2}a_{26}
$$

$$
\log(R1) = 0.89\log(a_{26})-10^{0.38} = \log(10^{-0.38}a_{26}^{0.89}) = \log(0.41\cdot a_{26}^{0.89})\Rightarrow R1\simeq \frac{a_{26}}{2}
$$

:::

@@

```{python}
#| echo: false
#| warning: false
logKLum_corr = round(compare_data(dt["logKLum"], dt["logL_K"], sigma = True)[3], 3)

@@

```{python}
#| label: fig-KLum-compare
#| fig-cap: "Comparison of the $L_K$ of the galaxies"
plt.close()
fig, ax = plt.subplots()
sns.regplot(x = dt["logKLum"], y = dt["logL_K"],
                 scatter_kws=dict(alpha=0.5, color='blue', edgecolors='white'),
                 line_kws=dict(alpha=0.7, color='red', linewidth=3))
temp = dt[["logKLum","logL_K"]].to_pandas().dropna()
results = scipy.stats.linregress(x = temp["logKLum"], y = temp["logL_K"])
plt.xlabel("log($L_{K, LCV} $"+"/{})".format(dt["KLum"].unit.to_string("latex")))
plt.ylabel("log($L_{K, HECATE} $"+"/{})".format(dt["KLum"].unit.to_string("latex")))
plt.legend(["Data",
            "$log(L_{K,HEC})=$" +f"{results.slope:.2f}"+"$\cdot log(L_{K,LCV})$"+
            f"${results.intercept:+.2f}$"+
            f"\n $R^2 = {results.rvalue**2*100:.2f}\%$"])
plt.show()
```

@@

So as we can see the usage of a linear fitting is not correct here.
But we can clearly see a linear correlation.

We will use the relative difference of the two luminosities, to see if statistically they are the same:

```{python}
#| echo: false
dt["diff_L_K"] = relative_diff(hec = dt["logL_K"].value, lcv = dt["logKLum"])
temp = dt[["logKLum", "logL_K", "diff_L_K"]].to_pandas()
sns.histplot(temp["diff_L_K"], kde = True)
plt.xlabel("Percentage change [%]")
plt.title("K-band Luminosities")
plt.show()
```

>>>>

In this script we will compare 2 catalogs, HECATE (@kovlakasHeraklionExtragalacticCatalogue2021) and LCV ([@karachentsevUPDATEDNEARBYGALAXY2013, @karachentsevSTARFORMATIONPROPERTIES2013a])

@@

```{python}
#| echo: false
import numpy as np
import matplotlib.pyplot as plt
from sklearn.metrics import r2_score
import astropy
from astropy.io import ascii
from astropy.coordinates import SkyCoord
from astropy.table import Table, join, QTable
import astropy.units as u
from astropy.visualization import quantity_support, hist
import scipy
from scipy.optimize import curve_fit
#quantity_support()
from astropy.stats import sigma_clip, SigmaClip
from astropy.modeling import models, fitting
from scipy.stats import pearsonr
import seaborn as sns
import pandas as pd
import glob
import os
from tabulate import tabulate
from IPython.display import Markdown, display

@@

Checking for completeness in galaxy catalogs is essential to ensure that the data accurately represents the true population of galaxies, for $z \approx 0$. Incomplete catalogs can lead to biased results in statistical studies, such as the distribution of galaxy luminosity, mass, or star formation rates. Additionally, missing galaxies, especially those at faint magnitudes or large distances, can distort cosmological measurements and hinder our understanding of galaxy formation and evolution.

@@

2.  Distances of $$
       D<11\ \text{Mpc}
       $$ {#eq-lcv-dis}

@@

### Difficulties in calculating the Completeness

As explained in (@karachentsevUPDATEDNEARBYGALAXY2013), it is difficult to calculate the completeness of the catalog.

1.  Completeness within a 10 Mpc radius is difficult to assess due to: - Variability in galaxy properties (luminosity, size, surface brightness, gas content) - Errors in distance measurements (Tully–Fisher method errors of \~20–25%), especially at the 10 Mpc boundary.
    -   Accurate distances are mainly known within \~5 Mpc.
    -   Non-Hubble motions (\~300 km/s)[^1] may make up half of the adopted velocity constraint (@eq-lcv-vel)
    -   Solution for our usage: only keep the galaxies inside a radius of $D=11 Mpc$
    -   HI sources in surveys with low angular resolution: "The presence around our Galaxy of hundreds of high-velocity clouds with low line-of-sight velocities and small W50 widths also provokes the inclusion of false “nearby” dwarf galaxies in the LV"(@karachentsevUPDATEDNEARBYGALAXY2013)
2.  Astro-Spam and Survey Errors: - Automatic surveys produce false detections (e.g., stars misclassified as galaxies, high-velocity clouds mistaken for dwarfs).
3.  Conditional Completeness Estimate: - Galaxies brighter than $M_B^c = -11^m$ or with linear diameters larger than $A_{26} = 1.0$ kpc show (40–60)% completeness. - "among the members of the Local Group (D \< 1 Mpc), only half of the galaxies have absolute magnitudes brighter than −11m. Consequently, more than half of the ultra-faint dwarf companions around normal galaxies, like the Sombrero galaxy (D = 9.3 Mpc), still remain outside our field of view."
4.  Undetected Ultra-Faint Dwarfs: - Many faint dwarf galaxies remain undetected beyond the Local Group. Estimated population of undetected dwarfs could be as large as $10^3–10^4$ within the LV.
5.  Surface Brightness Distribution: - Surface brightness remains consistent across distances, except for ultra-faint dwarfs ($SB <31 \text{mag}\cdot\text{arcsec}^{-2}$). - Faintest dwarf galaxies are detectable only nearby due to their resolution into individual stars.
6.  Luminosity-Size Relationship: - Observations align with cosmological models predicting $L\sim A^3$ (@navarroStructureColdDark1996), though deviations occur for extremely low surface brightness galaxies. - "The deviation from it at the extremely low surface brightness end is due to a systematic overestimation of dwarf galaxy sizes, the brightness profiles of which lie entirely below the Holmberg isophote."

[^1]: *Non-Hubble motion refers to the component of a galaxy's velocity that deviates from the uniform expansion of the universe, described by the equation* $V= H_0 d + v_p$,where $V$ is the total velocity, $H_0 d$ represents the Hubble flow, and $v_p$​ is the peculiar velocity. @Essay

@@

@@

@@

fig, (ax1, ax2) = plt.subplots(2, 1, sharex=True)
sns.histplot(main, bins = bin_edges, kde = True, label = "HECATE", ax = ax1)
sns.histplot(unique, bins = bin_edges, kde = True, label = "Unique for HECATE", ax = ax1)
ax1.set_title("Completenes of Distance for HECATE")
ax1.tick_params(labelbottom=True)
unique_pct = (unique_counts/main_counts)*100 
###
plt.subplot(2,1,2)
sns.lineplot(x = bin_edges[1:], y = unique_pct, marker = "o", ax = ax2)
#changing ylables ticks
# Set y-axis ticks for ax2 (completeness percentage)
ax2.set_yticks(np.arange(min(unique_pct), max(unique_pct) + 1, 5.0))

# Format the y-tick labels to display as percentages
y_value = ['{:,.0f}%'.format(x) for x in ax2.get_yticks()]
ax2.set_yticklabels(y_value)
ax2.set(xlabel = "Distance [{}]".format(lcv["Dis"].unit), ylabel="Completeness %")
plt.show()

@@

fig, (ax1, ax2) = plt.subplots(2, 1, sharex=True)
sns.histplot(main, bins = bin_edges, kde = True, label = "LCV", ax = ax1)
sns.histplot(unique, bins = bin_edges, kde = True, label = "Unique for LCV", ax = ax1)
ax1.legend()
ax1.set_title("Completenes of Distance for LCV")
ax1.tick_params(labelbottom=True)
unique_pct = (unique_counts/main_counts)*100 
###
sns.lineplot(x = bin_edges[1:], y = unique_pct, marker = "o", ax = ax2)

#changing ylables ticks
# Set y-axis ticks for ax2 (completeness percentage)
ax2.set_yticks(np.arange(min(unique_pct), max(unique_pct) + 1, 5.0))

# Format the y-tick labels to display as percentages
y_value = ['{:,.0f}%'.format(x) for x in ax2.get_yticks()]
ax2.set_yticklabels(y_value)
ax2.set(xlabel = "Distance [{}]".format(lcv["Dis"].unit), ylabel="Completeness %")
plt.show()

@@

fig, (ax1, ax2) = plt.subplots(2, 1, sharex=True)
sns.histplot(main, bins = bin_edges, kde = True, label = "HECATE",ax = ax1)
sns.histplot(unique, bins = bin_edges, kde = True, label = "Unique for HECATE",ax = ax1)
ax1.legend()
ax1.set_title("Completenes of Types for HECATE")
ax1.tick_params(labelbottom=True)
unique_pct = (unique_counts/main_counts)*100 
###

sns.lineplot(x = bin_edges[1:], y = unique_pct, marker = "o",ax = ax2)

#changing ylables ticks
# Set y-axis ticks for ax2 (completeness percentage)
ax2.set_yticks(np.arange(min(unique_pct), max(unique_pct) + 1, 5.0))

# Format the y-tick labels to display as percentages
y_value = ['{:,.0f}%'.format(x) for x in ax2.get_yticks()]
ax2.set_yticklabels(y_value)

ax2.set(xlabel = "Types of galaxies", ylabel="Completeness %")
plt.show()
######
#LCV
main = lcv["TType"].data
main = main[~np.isnan(main)]
unique = lcv_not_hec["TType"].data
unique = unique[~np.isnan(unique)]
counts, bin_edges = np.histogram(main)

@@

fig, (ax1, ax2) = plt.subplots(2, 1, sharex=True)
sns.histplot(main, bins = bin_edges, kde = True, label = "LCV", ax = ax1)
sns.histplot(unique, bins = bin_edges, kde = True, label = "Unique for LCV", ax = ax1)
ax1.legend()
ax1.set_title("Completenes of Types for LCV")
ax1.tick_params(labelbottom=True)
unique_pct = (unique_counts/main_counts)*100

sns.lineplot(x = bin_edges[1:] ,y = unique_pct, marker = "o", ax = ax2)

#changing ylables ticks
# Set y-axis ticks for ax2 (completeness percentage)
ax2.set_yticks(np.arange(min(unique_pct), max(unique_pct) + 1, 10.0))

# Format the y-tick labels to display as percentages
y_value = ['{:,.0f}%'.format(x) for x in ax2.get_yticks()]
ax2.set_yticklabels(y_value)

ax2.set(xlabel = "Types of galaxies", ylabel="Completeness %")
plt.show()
```

As we can see from the histograms (@fig-dis-comp) and (@fig-type-comp) the sample of unique galaxies of each catalog, gets smaller by an almost constant proportion (Inner join).

@@

For the comparison of our data we will mainly use two methods: linear regression and comparison of the distributions.

For the linear regression out main tools are scatter plots and $R^2$

1.  $R^2$: Measures the proportion of variance explained by the linear model.

2.  Slope of the Fitted Line: Should be close to 1 for a 1-1 correlation.

@@

4.  Plots: Plots are essential for visually assessing the relationship between two datasets, identifying correlations, trends, and outliers, and evaluating the fit of linear models. Beside scatter plots and histograms, another useful plot is the Correlation Heatmap.

    > Correlation Heatmaps: A correlation heatmap is a graphical tool that displays the correlation between multiple variables as a color-coded matrix. It's like a color chart that shows us how closely related different variables are. In a correlation heatmap, each variable is represented by a row and a column, and the cells show the correlation between them. The color of each cell represents the strength and direction of the correlation, with darker colors indicating stronger correlations.

@@

For the distrabution comparisons we will use histograms and Kernel Density Estimate (KDE) plots. The KDE plots visually represent the distribution of data, providing insights into its shape, central tendency, and spread.

Finally, we will examine the percentage change for each galaxy.

-   Percentage change: We can calculate the percentage change of the data for each galaxy and then we can see if the data are similar, based on minimum, the maximum and the mean value of the difference.

@@

| LCV | HECATE | Description |              Pearson Correlation \[-1,1\]              |
|:---------------:|:---------------:|:---------------:|:--------------------:|
| Dis |   D    |  Distance   | `{python} round(compare_data(dt["Dis"],dt["D"])[3],3)` |

@@

-   The average error of the distance in the HECATE catalog is $\overline{\delta D} = \pm 1.6$ Mpc, so the intercept is included in the error.
-   So we can assume that the Distances are the same

@@

-   The average error of the radial velocity in the HECATE catalog is $\overline{\delta V} = \pm 12\text{km}\cdot s^{-1}$ , so the intercept is included in the error.
-   So we can assume that the radial velocities are the same

@@

```{python}
#| label: fig-vel-pairplot
#| fig-cap: "The correlation Matrix of the Velocities. The lower left triangle is is composed of the scatter plots of the various velocities, the diagonal shows their distrubution and the upper right triangle shows the correlations of the Velocities" 
pairplot_with_correlation(vel_data)
```

@@

The strong correlation between the velocities reported by the HECATE and LCV catalogs can be attributed to the fact that both measure the same intrinsic galaxy velocities but reference them to different frames. For instance, velocities can be corrected for the movement of the Earth (heliocentric frame) or the for the Virgo-infall. These corrections shift the measured velocity slightly, but the fundamental measurement remains consistent. This explains why the velocities are highly correlated, as they represent the same physical phenomenon but adjusted for different reference points.

{{<pagebreak>}}

@@

"Morphological type of galaxy in the numerical code according to the classification by de Vaucouleurs et al. (1991). It should be noted that about three quarters of objects in the LV are dwarf galaxies, which require a more detailed morphological classification. For example, dwarf spheroidal galaxies and normal ellipticals are usually denoted by the same numerical code T \< 0, although their physical properties drastically differ. The classification problem arises as well for the “transient” type dwarf galaxies, T r, which combine the features of spheroidal (Sph) and irregular (Ir)systems. Due to small classification errors, such objects may “jump” from one end of the T scale to the other." (@karachentsevUPDATEDNEARBYGALAXY2013)

@@

types = dt[["T", "TType","E_T"]]
types["diff"] = dt["T"]-dt["TType"]
types = types[np.abs(types["diff"])<dt["TType"].std()]

@@

-   we can assume that the galaxies from the upper left and lower right regions of the plot are classified differently because of this
-   this might explain the large errors E_T of HECATE
-   why we see a difference in the distribution of (@fig-type-comp)

We can remove the "problematic" galaxies by only keeping the ones with: 
$$
|{T_{HECATE}-T_{LCV}}|<\sigma_{T_{LCV}} = 4.8
$$

-   The average uncertainty of the morphological type in the HECATE catalog is $\overline{\delta T} = \pm 1.4$, so the intercept is included in the error.
-   So we can assume that the Distances are the same

@@

```{python}
#| label: fig-incl-compare
#| layout-ncol: 2
#| layout-align: center
#| fig-cap: "Comparison of the Inclination of the galaxies"
#| fig-subcap: 
#| - "Distribution of the Inclination of the galaxies" 
#| - "Distribution of the Percentage Change" 
plt.close()
temp = dt[["inc", "INCL"]].to_pandas()

@@

temp["Percentage Change [%]"] = (-temp["inc"] + temp["INCL"])/temp["INCL"]
temp.loc[np.isinf(temp["Percentage Change [%]"]), "Percentage Change [%]"] = np.nan
sns.histplot(temp["Percentage Change [%]"], kde = True)
plt.show()
```

```{python}
temp.describe()
```

We can see that for values in the range $[\sim 30^\circ,\sim 80^\circ]$, the values of the LCV inclination are higher (@fig-incl-compare). However, since their means, median, min and maxes are similar and the percentage change is practically 0% (mean, median, $\sigma$ = 0 with a range $[-3\%,1\%]$), we can ignore the differences and assume they are the same values.

@@

it is not very clear if we truly have a correlation or not (@fig-axis-compare-1). We need to see the linear correlation of the decimal logarithms (@fig-axis-compare-2).

$\overline{R_1} = 3.9$ \[arcmin\], $\overline{a_{26}} = 9.3$ \[arcmin\], so the intercept is negligable.

$$
R_1 = 0.48\cdot a_{26}-0.34 \sim \frac{1}{2}a_{26}
$$

$$
\log(R1) = 0.89\log(a_{26})-10^{0.38} = \log(10^{-0.38}a_{26}^{0.89}) = \log(0.41\cdot a_{26}^{0.89})\Rightarrow R1\simeq \frac{a_{26}}{2}
$$
:::

@@

```{python}
#| echo: false
#| output: asis 
#| warning: false
logKLum_corr = round(compare_data(dt["logKLum"], dt["logL_K"], sigma = True)[3], 3)

@@

```{python}
#| label: fig-klum-compare
#| layout-ncol: 2 
#| fig-cap: "Comparison of the $L_K$ of the galaxies."
#| fig-subcap: 
#| - "Linear Regression with free paramaters, $y=ax+b$ "
#| - "Linear Regrassion $y=x$ "

plt.close()
fig, ax = plt.subplots()
sns.regplot(x = dt["logKLum"], y = dt["logL_K"],
                 scatter_kws=dict(alpha=0.5, color='blue', edgecolors='white'),
                 line_kws=dict(alpha=0.7, color='red', linewidth=3))
temp = dt[["logKLum","logL_K"]].to_pandas().dropna()
results = scipy.stats.linregress(x = temp["logKLum"], y = temp["logL_K"])
plt.xlabel("log($L_{K, LCV} $"+"/{})".format(dt["KLum"].unit.to_string("latex")))
plt.ylabel("log($L_{K, HECATE} $"+"/{})".format(dt["KLum"].unit.to_string("latex")))
plt.legend(["Data",
            "$log(L_{K,HEC})=$" +f"{results.slope:.2f}"+"$\cdot log(L_{K,LCV})$"+
            f"${results.intercept:+.2f}$"+
            f"\n $R^2 = {results.rvalue**2*100:.2f}\%$"])
plt.show()

data = dt[["logKLum", "logL_K"]]
data = dt[~np.isnan(data["logL_K"])]
logL_K_LCV = data['logKLum']  # Replace with actual column names
logL_K_HECATE = data['logL_K']  # Replace with actual column names

# Plot the data and both fits
plt.figure(figsize=(8, 6))
plt.scatter(logL_K_LCV, logL_K_HECATE, color='blue', label='Data', alpha =0.5, edgecolors="white")

# Plot free fit (with intercept)
plt.plot(logL_K_LCV, results.slope * logL_K_LCV + results.intercept, color='green', label='Free Fit: $log(L_{K,HEC})=$'+ f'{results.slope:.2f}'+ "$\cdot log(L_{K,LCV})$"+ f'{results.intercept:.2f}'+f"\n $R^2 = {results.rvalue**2*100:.2f}\%$")

slope_forced = 1
# Plot forced fit (through the origin)
results_f = scipy.stats.linregress(x = logL_K_HECATE, y = slope_forced*logL_K_LCV)
plt.plot(logL_K_LCV, slope_forced * logL_K_LCV, color='red', label='Forced Fit: $log(L_{K,HEC})=$' +f' = {slope_forced:.2f}'"$\cdot log(L_{K,LCV})$"+f"\n $R^2 = {results_f.rvalue**2*100:.2f}\%$")

plt.xlabel("log($L_{K, LCV} $"+"/{})".format(dt["KLum"].unit.to_string("latex")))
plt.ylabel("log($L_{K, HECATE} $"+"/{})".format(dt["KLum"].unit.to_string("latex")))
plt.legend()
plt.grid(True)
plt.show()
```

@@

To assess whether the two galaxy catalogs have a 1-1 correlation, I applied two types of linear fits to the data: a free fit and a forced fit. The free fit allows both the slope and intercept to vary freely, meaning the line can take on any slope and can cross the y-axis at any point. In contrast, for the forced fit, I set the slope to exactly 1 and fixed the line to pass through the origin, meaning the equation becomes y=xy=x. This forced fit represents a perfect 1-1 relationship, where the values from one catalog should match the other exactly, without any shift or scaling.
